Self Referencing Sequences

### 2014-10-26 draft.tex
edited iterative process

We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number. To see why this is true, we can simply give an algorithm to build the sequence. Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$.

\textbf{Algorithm:}
Assume that the sequence $\{s\}$ is initially just $a_i$. At the next step, add an additional $a_i -1$ repetitions of $a_i$ to the end of the sequence. Now if there is a second term in the sequence, read that term and write down that many copies of $a_{i+1}$. If there is no term, write down $a_{i+1}$.

NEED TO FINISH THIS.

Since we know that a self-referencing sequence is uniquely determined by the generating set and the starting number, we introduce the notation $s(A,a)$ to denote the unique sequence generated by set $A$ with starting number $a \in A$.

@@

One possible iterative process is as follows. Let the sequence $\{s^0\}$ be 2, and let the sequence $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$. This gives:
\begin{align*}
\{s^{(0)}\} &= 2\\
\{s^{(1)}\} &= 2,2\\
\{s^{(2)}\} &= 2,2,1,1\\
\{s^{(3)}\} &= 2,2,1,1,2,1\\
\{s^{(4)}\} &= 2,2,1,1,2,1,2,2,1\\
&\cdots
\end{align*}
Let $s^\infty$ be the limit of this process. We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
\begin{proof}
We show that this is true by simply showing that if $s^{(i)}$ is a prefix of
\end{proof}

>>>>

We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number. To see why this is true, we can simply give an algorithm to build the sequence. Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$. If we ever write down some $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
\begin{quote}
\textbf{Algorithm:}\\
Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that sequence. We read that second element, and then we add a block to the end of that sequence with length dictated by the value of the second element. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\

This procedure can continue deterministically so long as there is always an element to read. However, every time we write down a block, we read 1 element and write down at least 1 element, so the end of the sequence will always be after the spot we are reading. The only exception is when $a_i = 1$, but then since $a_{i+1} \neq 1$, this issue will never occur again.
\end{quote}

Thus, a self-referencing sequence is uniquely determined by the generating set and the starting number. From this point on, we will denote by $s(A,a)$ the unique sequence generated by set $A$ with starting number $a \in A$.

@@

One possible iterative process is as follows. Let the sequence $\{s^0\}$ be 2, and let the sequence $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$. This gives:
\begin{align*}
\{s^{(0)}\} &= 2\\
\{s^{(1)}\} &= 2,2\\
\{s^{(2)}\} &= 2,2,1,1\\
\{s^{(3)}\} &= 2,2,1,1,2,1\\
\{s^{(4)}\} &= 2,2,1,1,2,1,2,2,1\\
&\cdots
\end{align*}
Let $s^\infty$ be the limit of this process. We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
\begin{proof}
We prove this with induction. We show that if $s^{(i)}$ is a prefix of $s(\{1,2\},2)$, then $s^{(i+1)}$ is a longer prefix of $s(\{1,2\},2)$. The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(\{1,2\},2)$.

For the inductive step, observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence. We know that it describes a longer prefix because the prefix it describes has length equal to the sum of the original prefix. In other words, the length of $s^{(i+1)}$ must be equal to the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1). This holds true because $s^{(i)}$ merely contains a collection of lengths, which are summed up to give the next prefix. 
\end{proof}

It turns out that this same argument holds in general.

\begin{theorem}
Let $\{s^0\} = a$ for some $a \in A$ where $a \neq 1$, and let $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$. Then the limit of this process, $s^\infty$, is the sequence $S(A,a)$.
\end{theorem}

Note that we must specify that $a \neq 1$, because that is the only case in which the length of $s^{(i+1)}$ is not strictly greater than the length of $s^{(i)}$. The proof of this theorem is omitted, as it is essentially the same proof as for the $A = \{1,2\}, a = 2$ case.

It turns out that we can be less restrictive about what the starting sequence is. Going back to the case of $\{1,2\}$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2. For example, suppose $s^{(0)} = 2,1,1,1,2$. The rules for generating large $s^{(i)}$ give:
\begin{align*}
\{s^{(0)}\} &= 2,1,1,1,2\\
\{s^{(1)}\} &= 2,2,1,2,1,2,2\\
\{s^{(2)}\} &= 2,2,1,1,2,1,1,2,1,1,2,2\\
\{s^{(3)}\} &= 2,2,1,1,2,1,2,2,1,2,1,1,2,1,2,2,1,1\\
&\cdots
\end{align*}
